Domain-Driven, Generative Data Model for BigPetStore

### 2014-07-21 bdcloud_paper/latex/paper.tex
Spell checking:

Several synethic data set generators exist.  TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale. Targetted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.  Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}. Several approaches use domain-specific langugages \cite{Bruno2005,Hoag2007} to allow users to specify additional constraints such as which distributions to sample from or allow modeling basic relationships between fields such as simple arithmetic equations \cite{Alexandrov2012}. Arasu, et al. \cite{Arasu2011} demonstrated how constraint-solving techniques could be used as an alternative to procedural approaches. These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for analytics.

@@

\section{Overview of Model \& Simulation Procedures}
Our generatorive model combines various well-known mathematical modeling techniques to simulate the factors affecting customers' purchasing habits.  Each part of the model is derived from \emph{ab initio} assumptions.  In several cases, real data is used to parameterize parts of the model.

@@

For each proposed customer location zipcode $z$, we compute the distance between $z$ and every store's zipcode to find the closest store.  We compute the distance $d$ between $z$ and the zipcode $s_z$ of the closest store.  The weight $w_z$ for each zipcode $z$ is modeled using an exponential distribution with the average distance $\beta$.

@@

\subsection{Transactions}
A transaction model is created for each customer and integrated over the simulation's timeframe to generate transactions. Transaction records consist of a customer, a store, a date and time, and a list of items purchased and their prices.  The store is set to the store located closest to the customer. The date and time is given as a floating-point number with units of days since the start of the simulation.  (Timezones are not taken into account.)  Transaction times are generated from a Monte Carlo process based on the customer will use up items such as dog food (Section~\ref{sec:transaction-times}).  The purchased items are generated by a Hidden Markov Model (HMM) parameterized by the transaction time and amount of time remaining for the items in the customer's inventory (Section~\ref{sec:transaction-purchases}).

@@

where $a_i$ is the amount of item category $i$ ``stuff'' remaining at time $t$. The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$. The SDE is numerically integrated using the Euler-Maruyama method\cite{Klouden13} with timestep $\Delta t_n$ sampled from a Poisson process:

@@

where $\lambda_i$ is the item category's average usage rate, $\mu_i$ is the item category's average amount used per use, and $\sigma^2_i$ is the variance of the amount of the item category used per use. The parameter $\lambda_i$ is defined for each item cateogry.  The parameters $\mu_{i, c}$, and $\sigma^2_{i,c}$ are found based on a base rate for each item category multiplied by the number of pets of the appropriate species the customer has.  The exhaustion time $T_{E,i}$ for each item category is found by simulating the usage until $a_n = 0$.

The proposed transaction time is found from the exhaustion times.  For each item cateogry, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$. A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.  The earlist proposed transaction time is taken as the overall proposed transaction time $T_T$.

@@

After an item is chosen, the item category's exhaustion time is updated by propogating the item category usage SDE described in Section~\ref{sec:transaction-times}. \textcolor{red}{state transition, stop state}

@@

Secondly, we profiled our customers' purchasing habits to optimize the advertising campaign's effectivness by customizing the advertized products for each customer.  For each customer, we generated a feature vector by computing the frequency with which they would purchase the same brand or flavor from one transaction to the next.  We clustered the customers' feature vectors using the KMean algorithm with a range of cluster counts. Twenty clusters converged the error.

@@

Analysis of the feature vectors and clusters revealed four predominant purchasing profiles (Figure~\ref{fig:cluster_analysis}).  High frequencies of purchasing the same flavors repeatedly were tightly-correlated with high frequencies of purchasing the same brands -- these customers were likely to be happy with a particular item and kept purchasing the same item repeatedly.  For our advertising campaign, it would be unlikely to get these customers to purchase different items, so we should create incentives to purchase larger quantites, especially when inventory levels are high and needed to be depleted.  Other customers had a tendency to purchase either the same flavor or brand repeatedly, but varied in their choice of the other.  For customers who prefer a particular brand, we should target our advertising campaign to suggest other flavors sold by that brand.  Likewise, for customers who prefer a particular flavor, we should target our advertising campaign to suggest other brands with that flavor.  Lastly, some customers had very low frequencies of purchasing neither the same brand nr flavor.  Other factors, such as cost, not represented in this analysis may be driving these customers' purchasing habits and will need further study.

>>>>

Several synthetic data set generators exist.  TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale. Targeted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.  Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}. Several approaches use domain-specific languages \cite{Bruno2005,Hoag2007} to allow users to specify additional constraints such as which distributions to sample from or allow modeling basic relationships between fields such as simple arithmetic equations \cite{Alexandrov2012}. Arasu, et al. \cite{Arasu2011} demonstrated that constraint-solving techniques could be used as an alternative to procedural approaches. These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for analytics.

@@

\section{Overview of Model \& Simulation Procedures}
Our generative model combines various well-known mathematical modeling techniques to simulate the factors affecting customers' purchasing habits.  Each part of the model is derived from \emph{ab initio} assumptions.  In several cases, real data is used to parameterize parts of the model.

@@

For each proposed customer location zip code $z$, we compute the distance between $z$ and every store's zip code to find the closest store.  We compute the distance $d$ between $z$ and the zip code $s_z$ of the closest store.  The weight $w_z$ for each zip code $z$ is modeled using an exponential distribution with the average distance $\beta$.

@@

\subsection{Transactions}
A transaction model is created for each customer and integrated over the simulation's time frame to generate transactions. Transaction records consist of a customer, a store, a date and time, and a list of items purchased and their prices.  The store is set to the store located closest to the customer. The date and time is given as a floating-point number with units of days since the start of the simulation.  (Time zones are not taken into account.)  Transaction times are generated from a Monte Carlo process based on the customer will use up items such as dog food (Section~\ref{sec:transaction-times}).  The purchased items are generated by a Hidden Markov Model (HMM) parameterized by the transaction time and amount of time remaining for the items in the customer's inventory (Section~\ref{sec:transaction-purchases}).

@@

where $a_i$ is the amount of item category $i$ ``stuff'' remaining at time $t$. The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$. The SDE is numerically integrated using the Euler-Maruyama method\cite{Klouden13} with time step $\Delta t_n$ sampled from a Poisson process:

@@

where $\lambda_i$ is the item category's average usage rate, $\mu_i$ is the item category's average amount used per use, and $\sigma^2_i$ is the variance of the amount of the item category used per use. The parameter $\lambda_i$ is defined for each item category.  The parameters $\mu_{i, c}$, and $\sigma^2_{i,c}$ are found based on a base rate for each item category multiplied by the number of pets of the appropriate species the customer has.  The exhaustion time $T_{E,i}$ for each item category is found by simulating the usage until $a_n = 0$.

The proposed transaction time is found from the exhaustion times.  For each item category, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$. A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.  The earliest proposed transaction time is taken as the overall proposed transaction time $T_T$.

@@

After an item is chosen, the item category's exhaustion time is updated by propagating the item category usage SDE described in Section~\ref{sec:transaction-times}. \textcolor{red}{state transition, stop state}

@@

Secondly, we profiled our customers' purchasing habits to optimize the advertising campaign's effectiveness by customizing the advertised products for each customer.  For each customer, we generated a feature vector by computing the frequency with which they would purchase the same brand or flavor from one transaction to the next.  We clustered the customers' feature vectors using the KMeans algorithm with a range of cluster counts. Twenty clusters converged the error.

@@

Analysis of the feature vectors and clusters revealed four predominant purchasing profiles (Figure~\ref{fig:cluster_analysis}).  High frequencies of purchasing the same flavors repeatedly were tightly-correlated with high frequencies of purchasing the same brands -- these customers were likely to be happy with a particular item and kept purchasing the same item repeatedly.  For our advertising campaign, it would be unlikely to get these customers to purchase different items, so we should create incentives to purchase larger quantities, especially when inventory levels are high and needed to be depleted.  Other customers had a tendency to purchase either the same flavor or brand repeatedly, but varied in their choice of the other.  For customers who prefer a particular brand, we should target our advertising campaign to suggest other flavors sold by that brand.  Likewise, for customers who prefer a particular flavor, we should target our advertising campaign to suggest other brands with that flavor.  Lastly, some customers had very low frequencies of purchasing neither the same brand nor flavor.  Other factors, such as cost, not represented in this analysis may be driving these customers' purchasing habits and will need further study.
